---
title: "Longitudinal Data FACES, KUSS, PPMD"
author: "A.I. Georgevici"
date: "9 4 2020"
output: html_document

---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
Sys.setenv(LANG = "en")

library(tidyverse)
library(purrr)
library(readxl)
library(janitor)
library(gt)
library(here)
library(magrittr)
library(janitor)

path <- here("input/Rohdaten pro Patient")
```

```{r import the data}
longitudinal <- list.files(here("input/Rohdaten pro Patient"), pattern = ".xlsx") %>% 
  map_dfr(~ readxl::read_xlsx(
                path = paste0(path, "/", .x),
                range = cell_cols("A:Z"),
                col_types = "text" # import everyhting as text 
            ) %>%
        dplyr::select(-Time) %>% 
        janitor::remove_empty("rows") %>% 
        janitor::clean_names() %>% 
        rename(pseudo_id = pseudo, real_id = rea_i_id) %>% 
        drop_na(c(pseudo_id, real_id)) %>% # the rows without id are empty rows
        mutate(time = dplyr::row_number()) %>% 
        select(matches("id"), time, everything()) %>%  # the observations are in order
        mutate_at(vars(p1:interv), 
                  function(x) {
                    replace_na(x, 0) %>% # no entry represents 0
                    na_if("N") %>% # "N" represents Not_Available
                    as.integer() 
                  }
        ) 
  )
```

## Import the data

### Following assumptions:

* the rows are in chronological order, rows with no entry still have the time axis
* in the variables: "p1, p2, p3 ... interv" no entry represents a "0" and "N" represents N/A (not available)


### A quick birds-eye view of the column structure
```{r data structure 1}
longitudinal
```

### The percent of missing values per column. By study design, a patient has either FACES or KUSS values (never both)
```{r data structure 2}
library(DataExplorer)
DataExplorer::plot_missing(longitudinal, theme_config = list(legend.position = NULL)) # without legend
```

### A summary of the percentiles of numeric columns
```{r data structure 3}
library(skimr)
library(gt)
longitudinal %>% select_if(is.numeric) %>% 
  skim() %>% as_tibble() %>% select_all(~str_replace_all(., "skim_|numeric.", "")) %>%
  transmute(variable, missing_rate = 100*(1 - complete_rate), mean, sd, hist ) %>% 
  gt()
```

### The structure of KUSS

* kuss takes the max value from "kuss" or "sum of w, g, r, b, m"
* replace the "w,g,r,b,m" with NA if kuss is NA

```{r sync KUSS with sum of the sub-items}
psum <- function(...,na.rm=FALSE) { 
    rowSums(do.call(cbind,list(...)),na.rm=na.rm) } 

longitudinal %<>% mutate(kuss = pmax(kuss, psum(w,g,r,b,m, na.rm = FALSE), na.rm = TRUE)) %>%  
                  mutate_at(vars(w:m), ~ ifelse(is.na(kuss), NA, .)) 
```

In the missing_rate the percent of missing "w, g, r, b, m" when FACES is available !
```{r missing kuss sub_items}

longitudinal %>% drop_na(kuss) %>% select(w:m) %>% 
  skim() %>% as_tibble() %>% select_all(~str_replace_all(., "skim_|numeric.", "")) %>%
  transmute(variable, missing_rate = 100*(1 - complete_rate), mean, sd, hist ) %>% 
  gt()
```

The distribution of KUSS in percents of observations per sum of KUSS
```{r VIZ distribution of KUSS}
ggplot(
  data = count(longitudinal, kuss) %>%
    mutate(kuss = replace_na(kuss, "missing"), n = n/nrow(longitudinal))
  ) +
  geom_bar(aes(kuss, n), stat="identity") + 
  scale_y_continuous(labels = scales::percent_format(accuracy = 1))
```


The values in the original PPMD, there's a value "2" and "6".
Those values are corrected to 1. 
```{r the values in original ppmd}
longitudinal %>% pivot_longer(p1:p15) %>% distinct(name, value) %>% count(value)
longitudinal %<>% mutate_at(vars(p1:p15), ~ ifelse(. > 1, 1, .))
```

There are also missing values mixed with present values. Imputing 0 in those columns
At time1 (Aufwachraum) the PPMD is NA, imputing NA in those rows. 
```{r missing values mixed with present values in ppmd, sum of ppmd subitems}
longitudinal %>% mutate(nrNA = rowSums(is.na(select(., p1:p15)))) %>%
  filter(!nrNA %in% c(0,15)) %>% gt()

longitudinal %<>% mutate(nrNA = rowSums(is.na(select(., p1:p15)))) %>% 
  mutate_at(
    vars(p1:p15), 
    ~ ifelse(nrNA %in% c(1:14) & is.na(.), 0, .)
  ) %>% 
  mutate_at(
    vars(p1:p15), 
    ~ ifelse(time == 1, NA, .)
  ) %>% 
  select(-nrNA) %>% 
  mutate(
    ppmd = rowSums(select(., p1:p15)) 
  ) 
```

The distribution of PPMD in percents of observations per sum of ppmd
```{r VIZ distribution of PPMD}
ggplot(
  data = count(longitudinal, ppmd) %>%
    mutate(ppmd = replace_na(ppmd, "missing"), n = n/nrow(longitudinal))
  ) +
  geom_bar(aes(ppmd, n), stat="identity") + 
  scale_y_continuous(labels = scales::percent_format(accuracy = 1))
```

The number of observations per patient (pseudo_id)
```{r number of observations per patient}
longitudinal %>% count(pseudo_id) %>% count(n) %>% rename(nr_of_obs_per_pat = n, nr_of_pat = nn) %>% ggplot(.) + geom_bar(aes(nr_of_obs_per_pat, nr_of_pat), stat = "identity")
```

The number of interventions and non-interventions
```{r}
longitudinal %>% count(interv) %>% gt()
```

The number of observations (clustered per intervention) with the sum of FACES 
```{r}
longitudinal %>%
  mutate(
    sum_scales = rowSums(
      select(., kuss, faces, ppmd),
      na.rm = TRUE)
  ) %>%
  select(pseudo_id, time, kuss, faces, ppmd, sum_scales, interv) %>% 
  filter(sum_scales < 1) %>% 
  group_split(interv) %>% 
  map(~ head(.x, 20))
  
```

A dataset with a positive sum_scales (> 0)

```{r}
# the dataset with positive sum_scales(> 0)
positive <- longitudinal %>%  
  mutate(
    sum_scales = rowSums(
      select(., kuss, faces, ppmd),
      na.rm = TRUE)
  ) %>%
  filter(sum_scales > 0)

library(feather)
feather::write_feather(positive, path = here("input/positive.feather"))

positive 
```

Kuss Scale: n = nr. of children, interv (0/1) no intervention / intervention
```{r}
positive %>% group_by(interv) %>% count(kuss) %>% split(.$interv)
```

Faces Scale: n = nr. of children, interv (0/1) no intervention / intervention
```{r}
positive %>% group_by(interv) %>% count(faces) %>% split(.$interv)
```

PPMD Scale: n = nr. of children, interv (0/1) no intervention / intervention
```{r}
positive %>% group_by(interv) %>% count(ppmd) %>% split(.$interv)
```



`r longitudinal %>% filter(interv == 1) %>% pull(pseudo_id) %>% unique() %>% length()` children that had intervention from a total of `r longitudinal %>% pull(pseudo_id) %>% unique() %>% length()`.

Below the number of interventions per children

```{r}
longitudinal %>% filter(interv == 1) %>% count(pseudo_id) %>% arrange(desc(n)) %>% count(n) %>% rename(nr_of_interv_per_children = n, nr_of_children = nn) %>% gt()
```

