---
title: "production"
author: "A.I. Georgevici"
date: "8 4 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F)
```



I’ve been publishing screencasts demonstrating how to use the tidymodels framework, from first steps in modeling to how to tune more complex models. Today, I’m using this week’s #TidyTuesday dataset on beer production to show how to use bootstrap resampling to estimate model parameters.
```{r import data, message=FALSE}

library(tidyverse)
library(tidymodels)

brewing_materials_raw <- read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-03-31/brewing_materials.csv")

brewing_materials_raw %>%
  count(type, wt = month_current, sort = TRUE)
```


### How have some different brewing materials changed over time?
```{r VIZ evolution of materials over time}

brewing_filtered <- brewing_materials_raw %>%
  filter(
    type %in% c(
      "Malt and malt products",
      "Sugar and syrups",
      "Hops (dry)"
    ),
    year < 2016,
    !(month == 12 & year %in% 2014:2015)
  ) %>%
  mutate(
    date = paste0(year, "-", month, "-01"),
    date = lubridate::ymd(date)
  )

brewing_filtered %>%
  ggplot(aes(date, month_current, color = type)) +
  geom_point()


```



Below a linear model between the "malt products" and "sugar and syrups"
```{r VIZ linear models malt and sugar}

brewing_materials <- brewing_filtered %>%
  select(date, type, month_current) %>%
  pivot_wider(
    names_from = type,
    values_from = month_current
  ) %>%
  janitor::clean_names()

brewing_materials %>%
  ggplot(aes(malt_and_malt_products, sugar_and_syrups)) +
  geom_smooth(method = "lm") +
  geom_point()
```


In the next steps we build a linear model
```{r the linear model}

beer_fit <- lm(sugar_and_syrups ~ 0 + malt_and_malt_products,
  data = brewing_materials
)

summary(beer_fit)

tidy(beer_fit)

```


## Bootstrap resampling
```{r bootstrap resampling}
set.seed(123)
beer_bot <- bootstraps(brewing_materials, times = 1e3, apparent = TRUE)
```


We train a model to each of those bootstrap resamples.
```{r}
beer_models <- beer_bot %>% 
  mutate(
    model = map(splits, ~ lm(sugar_and_syrups ~ 0 + malt_and_malt_products, data = .)),
    coef_info = map(model, tidy)
  )
```


